---
title: "Introduction to R: Basics of R and R Data Structures" 
author: "Alex Lucchesi"    
date: "March 31st, 2023"
---

# Homework

1.  Create an array of two 3x3 matrices from the given vectors.

    ```{r}
    vec1 <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
    vec2 <- c(10, 11, 12, 13, 14, 15, 16, 17, 18)

    matrix_a <- matrix(vec1, nrow = 3, ncol=3)
    matrix_b <- matrix(vec2, nrow =3, ncol = 3)

    an_array <- array(c(matrix_a, matrix_b), dim=c(3,3,2))

    an_array
    ```

2.  In the above matrix that you created, print the value 5 by indexing in to the matrix.

    ```{r}
    an_array[2,2,1]

    ```

3.  Print the third row of the first matrix.

    ```{r}
    an_array[3,,1]
    ```

4.  Print the second column of both matrices.

    ```{r}
    an_array[,2,]

    ```

5.  Create two strings with your first and last name, respectively. Then concatenate them, and print the new value as "My name is \<your newly created concatenated variable will go here\>"

    ```{r}
    first_name <- "Sid"
    last_name <- "Hawker"
    full_name <- paste(first_name, last_name)

    sprintf('My name is %s', full_name)
    ```

6.  What values can a BOOLEAN contain?

    ```{r}

    y <- TRUE
    x <- FALSE

    class(y)
    is.logical(x)
    ```
