---
title: "Stats in R Assignment"
author: "Alex Lucchesi"
date: "2023/05/30"
output: html_document
---

## Assignment

For today's assignment, you will be given a blank notebook and a link to the [Call of Duty Dataset](https://www.kaggle.com/datasets/beridzeg45/call-of-duty-reviews) You will need to fill out the notebook and create your own code chunks to run code. Remember that we use this code without any quotation marks to create a code chunk: '`{r} '` \### Your Goal:

To receive credit for today's assignment, you will need to:

-   Import the dataset
-   Load the dataset into a tibble
-   Mutate a new column
-   Create a hypothesis regarding your data and communicate that hypothesis in a scientific method.
-   Create a notebook that has markdown cells, explaining your code and process/thoughts. **Your markdown should be well thought out and reflect a thorough analysis.**
-   Clean the data
-   Explore the data (include visualizations)
-   Create a statistical test and communicate the results

# I. Setup and Loading Data

```{r}
#loading necessary packages
library(tidyverse, dplyr)
```

```{r}
#using an underscore means we're making it a tibble
cod_df <- read_csv('Call of Duty Reviews.csv')

# view(cod_df)

```

# II. Cleaning Data

```{r}
rename_cols <- function(x){
  names(x) <- names(x) %>% str_replace_all(" ", '_') %>% tolower()
  return(x)
}
cod_df <- rename_cols(cod_df)
cod_tibble <- cod_df

View(cod_tibble)
```

```{r}
# checking for null values

sum(is.na(cod_tibble))

#removing them

clean_cod <- cod_tibble %>% drop_na()

View(clean_cod)

```

# III. Adjusting and mutating

```{r}
# mutating a column
clean_cod <- clean_cod %>%
  mutate(combo_score = ((critic_compound_score + user_compound_score)/2)*100) %>%
  mutate(platform_type = case_when(platform == "PC" ~ "PC",
                                   platform %in% c("X360", "XONE", "XBSX", "XBOX", "PS4", "PS3", "PS2")~ "MS/Sony",
                                   platform %in% c("WII", "WIIU", "GC")~"Nintendo",
                                   platform %in% c("DS", "VITA", "PSP")~"Handheld",
                                   platform=="iOS" ~ "Mobile")) 

```

# IV. Hypotheses & Linear Model

### Null: COD reviews are steady over time.

### Alternative Hypothesis: There reviews for COD games are getting better (or worse) over time

Method: Dependent variable: Metacritic Score, Independent Variable: Year of Release

Based on the summary of the linear model, there does not seem to be a statistical trend of game ratings going up or down over time.

Let's go to a visualization to check!

```{r}
fit <- lm(clean_cod$metascore ~ clean_cod$release_year)

summary(fit)
```

# V. is for Visuals

```{r}
ggplot(clean_cod, aes(release_year, metascore, color=platform, shape=platform_type))+
  geom_point(size=4)
```

# X. Other Viz

```{r}
avg_player_rating <- mean(clean_cod$user_compound_score)

ggplot(clean_cod) +
  geom_histogram(binwidth = .1, aes(x=user_compound_score, alpha=0.7), fill ="blue", color = 'black')+
  geom_vline(xintercept = avg_player_rating, color = "white", linetype = 'dashed') +
  labs(title = 'distribution of user ratings', x='user rating', y= 'frequency')
```

```{r}

avg_critic_rating <- mean(clean_cod$critic_compound_score)

ggplot(clean_cod) +
  geom_histogram(binwidth = .1, aes(x=critic_compound_score, alpha = 0.7), fill ="orange", color = 'black')+
  geom_vline(xintercept = avg_critic_rating, color = "white", linetype = 'dashed') +
  labs(ttile = 'distribution of user ratings', x='user rating', y= 'frequency')
```

```{r}
ggplot(clean_cod) +
  geom_histogram(binwidth = .1, aes(x=critic_compound_score, alpha = 0.7), fill ="orange", color = 'black')+
  geom_histogram(binwidth = .1, aes(x=user_compound_score, alpha=0.7), fill ="blue", color = 'black')
```

```{r}

```
