---
title: "R for Data Exploration, Visualizations, Functions, Conditionals, Loops, Tibbles, and Tidy Data" 
author: "Alex Lucchesi"    
date: "March 31st, 2023"
editor_options: 
  markdown: 
    wrap: 72
---

# **R Libraries: ggplot2, dplyr**

### Homework

1.  Do 5 CodeWars problems in R. Share your solutions here:

    <https://www.codewars.com/kata/53af2b8861023f1d88000832/train/r>

    ```{r}
    are_you_playing_banjo <- function(name){
     if(substr(name, 1,1)=="R" | substr(name, 1,1)=="r"){
       return(paste(name, "plays banjo"))
     }else{
       return(paste(name, "does not play banjo"))
     }
    }
    ```

    <https://www.codewars.com/kata/57cc975ed542d3148f00015b/train/r>

    ```{r}
    check <- function(a, x){
     for(i in a){
       if(i == x){
         return(TRUE)
       }
     }
      return(FALSE)
    }
    ```

    <https://www.codewars.com/kata/5875b200d520904a04000003/train/r>

    ```{r}
    enough <- function(cap, on, wait){
      if(on+wait <= cap){
        return(0)
      }else{
        return((on+wait)-cap)
      }
    }
    ```

    <https://www.codewars.com/kata/568dcc3c7f12767a62000038/train/r>

    ```{r}
    set_alarm <- function(employed, vacation) {
      if(employed==FALSE){FALSE}
      else if(vacation==FALSE){TRUE}
      else{FALSE}
    }
    ```

    <https://www.codewars.com/kata/5a6663e9fd56cb5ab800008b/train/r>

    ```{r}
    human_years_cat_years_dog_years <- function(human_years){
      cat <- 15
      dog <- 15
      if (human_years == 1){c(human_years, cat, dog)}
      else if(human_years == 2){c(human_years, cat+9, dog+9)}
      else{
        cat <- cat + 9
        dog <- dog + 9
        c(human_years, cat+((human_years-2)*4), dog+((human_years-2)*5))
      }
    }
    ```

2.  Create a function that checks an input of type `int` and returns
    fizz if the number is a multiple of 3, buzz if the number is a
    multiple of 5, and fizzbuzz if the number is a multiple of 3 and 5.

    ```{r}
    fizzbuzz <- function(y){
      if(y%%15==0){
        return('fizzbuzz')
      }else if(y%%3==0){
        return('fizz')
      }else if(y%%5==0){
        return('buzz')
        }
    }

    print(fizzbuzz(48))
    ```

3.  **Mario Challenge:** In the classic Super Mario games, Mario ends
    each level by ascending a flight of stairs and leaping off the top
    to arrive at the castle at the end. (See
    [here](https://youtu.be/-avspZlbOWU?t=55) for details). We can
    imagine that each of the bricks of the stairs is a character X.
    Create a function that prints out a set of stairs. For example:

    If the function is called marioStairs() and we pass in the value 4,
    we would expect to see stairs with a length of 4 and a height of 4:

    X

    XX

    XXX

    XXXX

    ```{r}
    marioStairs<- function(a){
      for(i in 1:a){
        print(paste(rep("X",i),collapse=""))
      }
    }

    marioStairs(8)
    ```

4.  Find a built-in dataset you would like to explore. Explore the
    dataset and use the cheat sheet you built in Week 1 to label
    continuous and discrete data. Analyze the dataset using dplyr,
    create graphs, mutate two columns, and formulate a hypothesis about
    your data.

#### **Data Types:**

-   sleep_total: Continuous

-   awake: continuous

```{r}
library(ggplot2)
# view(msleep)
```

```{r}
print(filter(msleep, vore=="omni"))
print(select(msleep, contains("wt")))

msleep %>%
  filter(order > "M")%>%
  arrange(order)
```

```{r}
enhanced_msleep<- msleep%>%
  mutate(total_hours = sleep_total + awake,
         brain_perc = (brainwt / bodywt)*100)

view(enhanced_msleep)
```

```{r}
ggplot(enhanced_msleep, aes(sleep_rem, brain_perc, color = order))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()
```
