---
title: "Dataset Capstone"
output: html_notebook
---

# Overview

## Part 2: 2-D plots

# R

## Preparing the Data

### Imports

```{r}

library(tidyverse)
library(plotly)

exif <- read_csv("capstone_exif.csv")
img_data <- read_csv("capstone_img_data.csv")

```

### Cleaning

**Pre Import:**\
Sub-image data for main images (0) was bugged in the first hours of image processing. This was fixed in Excel during the data collation stage.

[Editors Note: numeric values looked consistent in Excel and R Studio, but were interpreted as characters, necessitating a large amount of manual re-typing in part 2. At the time it seemed faster than backing up and figuring out the 'right' way to fix them]

**EXIF:**

-   Fix column headers
-   Drop columns: date_time_original, modify_date, lens_info, f_number, focal_length
-   NA values - JFIF \<- 0,  subject_area \<- \>depends on data type\< "0 0 0 0"

```{r}
names(exif) <- gsub("([a-z0-9])([A-Z])", "\\1_\\2", names(exif))
names(exif) <- names(exif) %>% tolower()

exif_tidy <- select(exif, -c(date_time_original, modify_date, lens_info, fnumber, focal_length))
exif_tidy <- replace_na(exif_tidy, list(subject_area = "0 0 0 0", jfifversion = 0))

```

**img_data:**

-   Drop columns: flickr, img_loc, the_image, crop_coords, r_mode, b_mode, g_mode, img_height, img_width, do_img_at
-   Na values - post_2-6_hsl \<- (-1,-1,-1)

```{r}
imgsd_tidy <- select(img_data, -c(flickr, img_loc, the_image, img_width, img_height, crop_coords, do_img_at, r_mode, b_mode, g_mode))

imgsd_tidy <- replace_na(imgsd_tidy, list(
  post_2_hsl = "(-1, -1, -1)",
  post_3_hsl = "(-1, -1, -1)",
  post_4_hsl = "(-1, -1, -1)",
  post_5_hsl = "(-1, -1, -1)",
  post_6_hsl = "(-1, -1, -1)"
  )
  )

```

### Basic Feature Engineering

**EXIF**

-   Split date_time_original to year, month-day, time columns
-   Date column uses a placeholder year so month-to-month comparisons are consistent

```{r}

exif_tidy <- exif_tidy %>% separate(create_date, into = c('date', 'time'), sep = " ", remove = TRUE) %>% separate(date, into = c('year', 'month', 'day'), sep = ":") 

exif_tidy$date <- as.Date(paste("1881", exif_tidy$month, exif_tidy$day, sep = "-"), format ="%Y-%m-%d")
```

**Img_data**

-   Count by flickr id

    -   Flag counts under 10
    -   Flag counts under 6 (ie: subimage 5 not available)
    -   Output list of all flickrIDs with less than 10 records for additional (future) processing

```{r}

subimg_qty <- imgsd_tidy %>% count(using_id)

```

To my (happy) surprise, only 5 images have less than 10 results and only 2 have less than 6. In the interest of time, I'm noting these IDs by hand and simply removing them from my working data

```{r}

good_ids <- subimg_qty[subimg_qty$n >=6, "using_id"]

imgsd_tidy <- imgsd_tidy %>% filter(using_id %in% good_ids$using_id)

exif_tidy <- exif_tidy %>% filter(flickr_id %in% good_ids$using_id)

```

-   Total pixels = full\_(hue)\_count(s) (dimension data incorrect for first 1400 records)

```{r}

imgsd_tidy <- imgsd_tidy %>% mutate(total_pixels = full_red_count +
                                      full_orange_count +
                                      full_yellow_count +
                                      full_green_count +
                                      full_cyan_count +
                                      full_blue_count + 
                                      full_purple_count +
                                      full_mag_count)

```

-   Split pixel lists/tuples for 3-d mapping

```{r}
# remove brackets from string-encapsulated lists
imgsd_tidy$center_hsl <- str_replace(imgsd_tidy$center_hsl, '\\[|\\]', '')

# remove parens from string-encapsulated tuples
imgsd_tidy <- imgsd_tidy %>% mutate_all(~ gsub('\\(|\\)', '', .))

# separate pixel values into individual columns
imgsd_split <- imgsd_tidy %>% 
  separate(center_rgb, 
           into = c('center_r', 'center_g', 'center_b'), 
           sep = ',') %>%
  separate(post_top_hsl, 
           into = c('post_top_hue', 'post_top_sat', 'post_top_light'), 
           sep = ',') %>%
  separate(post_2_hsl, 
           into = c('post_2_hue', 'post_2_sat', 'post_2_light'), 
           sep = ',') %>%
  separate(post_3_hsl, 
           into = c('post_3_hue', 'post_3_sat', 'post_3_light'), 
           sep = ',') %>%  
  separate(post_4_hsl, 
           into = c('post_4_hue', 'post_4_sat', 'post_4_light'), 
           sep = ',') %>%
  separate(post_5_hsl, 
           into = c('post_5_hue', 'post_5_sat', 'post_5_light'), 
           sep = ',') %>%
  separate(post_6_hsl, 
           into = c('post_6_hue', 'post_6_sat', 'post_6_light'), 
           sep = ',') %>%
  separate(center_hsl, 
           into = c('center_hue', 'center_sat', 'center_light'), 
           sep = ',') %>%
  separate(common_hsl_1_val, 
           into = c('common_hsl_1_hue', 'common_hsl_1_sat', 'common_hsl_1_light'), 
           sep = ',') %>%
  separate(common_hsl_2_val, 
           into = c('common_hsl_2_hue', 'common_hsl_2_sat', 'common_hsl_2_light'), 
           sep = ',') %>%
  separate(common_hsl_3_val, 
           into = c('common_hsl_3_hue', 'common_hsl_3_sat', 'common_hsl_3_light'), 
           sep = ',') %>%
  separate(common_hsl_4_val, 
           into = c('common_hsl_4_hue', 'common_hsl_4_sat', 'common_hsl_4_light'), 
           sep = ',')

```

... I probably should have saved those independently during the python image processing stage.

-   To be added as needed:

    -   Using total_pixels to calculate specific ratios

## Exploration

### 2D Scatter

```{r}
custom_colors = c('#de4bcd','#8f2411', '#f5b049', '#9c8a1c', '#b3e32d',  '#1a9615', '#05ffb4','#055eab', '#6b63ff','#45008f')

imgsd_split$common_hsl_4_hue <- as.numeric(imgsd_split$common_hsl_4_hue)
imgsd_split$common_hsl_4_sat <- as.numeric(imgsd_split$common_hsl_4_sat)

fig <- plot_ly(imgsd_split, 
               x=~common_hsl_4_hue, 
               y= ~common_hsl_4_sat, 
               type = "scatter", mode="markers", size = 2, color = ~sub_img, colors = custom_colors)

fig

```

Observation: Here we can more closely examine the hue trends present in the images. I'm fascinated by the cluster at 105 hue/60-80 saturation which is present in many sub-images but not 3, 9, or 6.

-   Light min x light max

```{r}

imgsd_split$light_min_val <- as.numeric(imgsd_split$light_min_val)
imgsd_split$light_max_val <- as.numeric(imgsd_split$light_max_val)

fig <- plot_ly(imgsd_split, 
               x=~light_min_val, 
               y= ~light_max_val, 
               type = "scatter", mode="markers", 
               size = 2, color = ~sub_img, colors = custom_colors)

fig

```

-   Gen bright x gen dark

```{r}
imgsd_split$gen_bright_count <- as.integer(imgsd_split$gen_bright_count)


imgsd_split$gen_dark_count <- as.integer(imgsd_split$gen_dark_count)

fig <- plot_ly(imgsd_split, 
               x=~gen_bright_count, 
               y= ~gen_dark_count, 
               type = "scatter", mode="markers", 
               size = 2, color = ~sub_img, colors = custom_colors)

fig
```

-   Sat min x sat max

```{r}

imgsd_split$sat_min_val <- as.numeric(imgsd_split$sat_min_val)
imgsd_split$sat_max_val <- as.numeric(imgsd_split$sat_max_val)

fig <- plot_ly(imgsd_split, 
               x=~sat_min_val, 
               y= ~sat_max_val, 
               type = "scatter", mode="markers", 
               size = 2, color = ~sub_img, colors = custom_colors)

fig

```
